Untitled

### 2020-10-07 ms/manuscript.Rmd
Update MS

---
title: "Untitled"
output: html_document
---

@@

The movement of plants from water to land was one of the most important evolutionary events in the history of the earth, fundamentally altering the global carbon and water cycle and setting the stage for modern terrestrial ecosystems (Graham, 1993; Kenrick and Crane, 1997).
The transition from aquatic to terrestrial growth was complex, requiring the evolution of a suite of adaptations to survive in dry habitats bombarded with radiation (Waters, 2003).
The earliest land plants were poikilohydric, i.e., they lacked any mechanisms to regulate water relations (Raven, 1999).
Therefore, one particularly important early adaptation that enabled plants to survive on land was desiccation tolerance (DT), the ability to lose water until reaching equilibrium with atmospheric humidity, then recover upon rewetting (Mishler and Churchill, 1985).

Compared with the ubiquitous DT among the earliest land plants, the occurrence of DT varies across extant land plant lineages and life stages.
Many bryophytes have retained vegetative DT in the gametophyte stage (Oliver et al. 2005, and references therein).
In extant vascular plants (tracheophytes), DT occurs in certain non-vegetative stages (e.g., seeds and pollen), but has largely been lost in vegetative tissues (with some notable exceptions, the so-called 'resurrection plants').
Instead, tracheophytes avoid desiccation with adaptations including a waxy cuticle, stomata, and an extensive vascular system.
Correlated with these differences in functional strategies to cope with water stress is a major difference in life cycle: in most tracheophytes (including all seed plants) the long-lived, dominant phase of the life cycle is the sporophyte (diploid phase), whereas in bryophytes it is the gametophyte (haploid phase).
Little is known about the details of the transition from desiccation tolerance to sensitivity, nor how this took place in conjunction with the switch from a gametophyte to a sporophyte-dominated life cycle; yet, these two steps underpin the radiation of plants onto dry land.

Ferns represent an important, understudied group of tracheophytes for investigating the transition from desiccation tolerance to desiccation avoidance in land plants.
Ferns are the only major lineage among land plants with sporophytes and gametophytes that are capable of growing independently from each other.
These two life stages differ drastically in their morphology and physiology---fern sporophytes have true vascular tissue and are homoiohydric (i.e., they regulate their internal water content), whereas fern gametophytes are much smaller (often *\<*1 cm), lack vascular tissue, and are poikilohydric.
Recent studies indicate that the gametophytes of many fern species are desiccation tolerant, while sporophytes of the same species are thought to lack DT (Watkins et al., 2007; Pittermann et al., 2013).
Thus, ferns present a unique opportunity to observe the transition from desiccation tolerance to desiccation avoidance across independently growing life phases within a single organism.

@@

***Field survey***---We surveyed filmy ferns on Moorea as part of a larger study including all ferns of Moorea and Tahiti (Nitta et al., 2016).
During this study, 17 sampling sites were established from ca. 200 m to 1200 m primarily along trails to the summits of three large peaks on Moorea: Mt. Rotui (899 m), Mt. Mouaputa (880 m), and Mt. Tohiea (1206 m; Fig. 2).
At each site, sporophytes were sampled in a single 10 m *×* 10 m plot, and gametophytes sampled in two to three 50 cm *×* 50 cm subplots located within the sporophyte plot.
Epiphytes were sampled to a maximum height of ca. 2 m on their host plants.
Unlike many other fern species, filmy fern gametophytes have unique morphologies that allow them to be identified to family (e.g., filamentous, ribbon with round gemmae; Nayar and Kaur 1971).
This allowed us to specifically use field-collected filmy fern gametophytes for physiological tests, which we later identified to species using DNA barcoding (see below).
We characterized the elevational range of each species based on the results of the field survey; any species found with gametophytes occurring *\>*200 m above the maximum or below the minimum observed elevational range of sporophytes was considered 'widespread' (i.e., facultatively independent sensu Rumsey and Sheffield \[1996\]).
We scored growth habit of each species based on field observations as terrestrial, saxicolous, low elevation epiphyte (mostly occurring outside of cloud forest, i.e., below ca. 500 m), or high elevation epiphyte (mostly occurring in cloud forest, i.e., above ca. 500 m).

@@

We used general linear mixed models (GLMMs) to investigate the effect of growth habit, range size, and life stage on DT (% recovery of F~v~/F~m~ after 48 h recovery from desiccation at --86 MPa for 2 d) and photosynthetic parameters (ETR~max~ and PPFD~95%~) using the MCMCglmm R package (Hadfield, 2010).
For each of the three response variables (DT, ETR~max~ , and PPFD~95%~), we used growth habit (epiphyte vs. non-epiphyte \[i.e., terrestrial or saxicolous\]) or range size (widespread vs. not widespread) and generation (sporophyte vs. gametophyte) and their interactions as fixed effects.
Two species were observed both as low-elevation epiphytes and saxicoles (Table 1), but we treated them as epiphytic to distinguish them from exclusively saxicolous species, which we regard as the more specialized growth habit.
We accounted for intraspecific variation and phylogeny (only for response variables with significant phylogenetic signal) by including species (within generation) and an inverse phylogenetic variance-covariance matrix as random effects, respectively (Hadfield, 2010).
We used an inverse-Gamma distribution for all priors, and ran analyses for 500 000 iterations, with burnin after 1000 iterations and thinning every 500 iterations.
Comparisons of DT, ETR~max~ , and PPFD~95%~ between gametophytes and sporophytes of the same species do not involve phylogeny, so we used a *t*-test for these.

>>>>

---
title: "Untitled"
output: html_document
bibliography: references.bib
editor_options: 
  markdown: 
    wrap: sentence
---

@@

The movement of plants from water to land was one of the most important evolutionary events in the history of the earth, fundamentally altering the global carbon and water cycle and setting the stage for modern terrestrial ecosystems [@Graham1993; @Kenrick1997a].
The transition from aquatic to terrestrial growth was complex, requiring the evolution of a suite of adaptations to survive in dry habitats bombarded with radiation [@Waters2003].
The earliest land plants were poikilohydric, i.e., they lacked any mechanisms to regulate water relations [@Raven1999].
Therefore, one particularly important early adaptation that enabled plants to survive on land was desiccation tolerance (DT), the ability to lose water until reaching equilibrium with atmospheric humidity, then recover upon rewetting [@Mishler1985].

Compared with the ubiquitous DT among the earliest land plants, the occurrence of DT varies across extant land plant lineages and life stages.
Many bryophytes have retained vegetative DT in the gametophyte stage [@Oliver2005, and references therein].
In extant vascular plants (tracheophytes), DT occurs in certain non-vegetative stages (e.g., seeds and pollen), but has largely been lost in vegetative tissues (with some notable exceptions, the so-called 'resurrection plants').
Instead, tracheophytes avoid desiccation with adaptations including a waxy cuticle, stomata, and an extensive vascular system.
Correlated with these differences in functional strategies to cope with water stress is a major difference in life cycle: in most tracheophytes (including all seed plants) the long-lived, dominant phase of the life cycle is the sporophyte (diploid phase), whereas in bryophytes it is the gametophyte (haploid phase).
Little is known about the details of the transition from desiccation tolerance to sensitivity, nor how this took place in conjunction with the switch from a gametophyte to a sporophyte-dominated life cycle; yet, these two steps underpin the radiation of plants onto dry land.

Ferns represent an important, understudied group of tracheophytes for investigating the transition from desiccation tolerance to desiccation avoidance in land plants.
Ferns are the only major lineage among land plants with sporophytes and gametophytes that are capable of growing independently from each other.
These two life stages differ drastically in their morphology and physiology---fern sporophytes have true vascular tissue and are homoiohydric (i.e., they regulate their internal water content), whereas fern gametophytes are much smaller (often *\<*1 cm), lack vascular tissue, and are poikilohydric.
Recent studies indicate that the gametophytes of many fern species are desiccation tolerant, while sporophytes of the same species are thought to lack DT [@WatkinsJr.2007c; @Pittermann2013a].
Thus, ferns present a unique opportunity to observe the transition from desiccation tolerance to desiccation avoidance across independently growing life phases within a single organism.

@@

***Field survey***---We surveyed filmy ferns on Moorea as part of a larger study including all ferns of Moorea and Tahiti (Nitta et al., 2016).
During this study, 17 sampling sites were established from ca.
200 m to 1200 m primarily along trails to the summits of three large peaks on Moorea: Mt. Rotui (899 m), Mt. Mouaputa (880 m), and Mt. Tohiea (1206 m; Fig. 2).
At each site, sporophytes were sampled in a single 10 m *×* 10 m plot, and gametophytes sampled in two to three 50 cm *×* 50 cm subplots located within the sporophyte plot.
Epiphytes were sampled to a maximum height of ca.
2 m on their host plants.
Unlike many other fern species, filmy fern gametophytes have unique morphologies that allow them to be identified to family (e.g., filamentous, ribbon with round gemmae; Nayar and Kaur 1971).
This allowed us to specifically use field-collected filmy fern gametophytes for physiological tests, which we later identified to species using DNA barcoding (see below).
We characterized the elevational range of each species based on the results of the field survey; any species found with gametophytes occurring *\>*200 m above the maximum or below the minimum observed elevational range of sporophytes was considered 'widespread' (i.e., facultatively independent sensu Rumsey and Sheffield $$1996$$).
We scored growth habit of each species based on field observations as terrestrial, saxicolous, low elevation epiphyte (mostly occurring outside of cloud forest, i.e., below ca. 500 m), or high elevation epiphyte (mostly occurring in cloud forest, i.e., above ca. 500 m).

@@

We used general linear mixed models (GLMMs) to investigate the effect of growth habit, range size, and life stage on DT (% recovery of F~v~/F~m~ after 48 h recovery from desiccation at --86 MPa for 2 d) and photosynthetic parameters (ETR~max~ and PPFD~95%~) using the MCMCglmm R package (Hadfield, 2010).
For each of the three response variables (DT, ETR~max~ , and PPFD~95%~), we used growth habit (epiphyte vs. non-epiphyte $$i.e., terrestrial or saxicolous$$) or range size (widespread vs. not widespread) and generation (sporophyte vs. gametophyte) and their interactions as fixed effects.
Two species were observed both as low-elevation epiphytes and saxicoles (Table 1), but we treated them as epiphytic to distinguish them from exclusively saxicolous species, which we regard as the more specialized growth habit.
We accounted for intraspecific variation and phylogeny (only for response variables with significant phylogenetic signal) by including species (within generation) and an inverse phylogenetic variance-covariance matrix as random effects, respectively (Hadfield, 2010).
We used an inverse-Gamma distribution for all priors, and ran analyses for 500 000 iterations, with burnin after 1000 iterations and thinning every 500 iterations.
Comparisons of DT, ETR~max~ , and PPFD~95%~ between gametophytes and sporophytes of the same species do not involve phylogeny, so we used a *t*-test for these.
